##  Final Project: Data Analysis and Preliminary Analytics
### Team 3 :
#### Sindhu Bhattarai
#### Emina Belekanich
#### Travis Lloyd

**Data Importing and Pre-processing:**
In this part we are going to first get necessary package,import our csv file, start checking for NAs and outliers.

                                    
```{r}
library(pacman)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
library(lubridate)
library(rgl)
require(scatterplot3d)


data <- read.csv("house_sales.csv")

attach(data)
head(data)
```


Converting date column using lubridate package

```{r}
data$date <- parse_date_time(data[,2], orders = "%Y%m%%d%H%M%S")
head(data)
```

Checking where empty values are located.

```{r}
colSums(is.na(data))
```

fill in sqft_living with basic addition of sqft_above and sqft_basement

```{r}
data$sqft_living <- data$sqft_above + data$sqft_basement
head(data)
```
We have filled in sqft_living  by adding sqft_above and sqft_basement.

Fill in bedrooms and bathrooms
```{r}
data$bedrooms[which(is.na(data$bedrooms))] <- 2.5
data$bathrooms[which(is.na(data$bathrooms))] <- 2
data$sqft_lot[which(is.na(data$sqft_lot))] <- 6500
```
Here we have filled the NAs with median value of the respective vector.

**Data Analysis and Visualization:**

In this part we are going to visualize the outliers using  boxplots, deal with outliers and check the correlations between the variables to better know the relationship among the variables. 


Removing date column to check correlation table

```{r}
data.cor <- subset(data, select = -c(date))
head(data.cor)
cor(data.cor)
```

Columns with high correlation values with grade as dependent variable:price, bathrooms, sqft_living, floors,sqft_above, yr_built, sqft_living15

```{r}
data2 <- subset(data, select = c(grade, price, bathrooms, sqft_living, floors, yr_built, sqft_living15))
head(data2)
colSums(is.na(data2))
```

missing data is filled in.
Moving forward we will look at the outliers of the main categories that we will be inspecting.

```{r}
boxplot(data2$bathrooms)

boxplot(data2$floors)

boxplot(data2$yr_built)

boxplot(data2$sqft_living)
boxplot(data2$sqft_living15)

boxplot(data2$price)

boxplot(data2$grade)
```

bathroom outlier > 3.5
floors no outliers
yr_built no outliers
sqft_living outliers > 3500
sqft_living_15 outliers > 3750
Price outliers > 900000

Removing outliers and  subsetting necessary variables

```{r}
data2 <- subset(data2, sqft_living15 < 3750
                       & sqft_living < 3500 
                       & bathrooms < 3.5 
                       & price <= 900000)
```

**Data Analytics:**
In this part we are going to construct a regression model to make the prediction. 


Making a scaled version of data2 to check the correlations use in linear regression models                 
```{r}
data2.scaled <- scale(data2)
cor(data2.scaled)
data2.df <- as.data.frame(data2.scaled)
```
 We noticed  in the correlation table that grade has comparatively higher correlation with price, sqft_living, yr_built and sqft_living15. Here in our first regression model we have taken grade as the dependent variable. 
 
 Second we noticed that price also has  comparatively higher correlation with grade, sqft_living and sqft_living15. In our second regression model we have taken price as the dependent variable and rest other variables with higher correlations as predictor variables. 

Making a multiple linear regression model of both grade and price and check which one is more relevant. 

```{r}
m1 <- lm(grade ~ price + sqft_living + yr_built + sqft_living15, data = data2)
m2 <- lm(price ~ grade + sqft_living  + sqft_living15, data = data2)
summary(m1)
summary(m2)
```

The equation of multiple regression line is y= m1*x1 + m2*x2 + m3*x3 +.... mn*xn + c

Here as we can see in our first regression summary. The equation of the regression line of grade prediction is: 
grade = (1.902e-06 * price) +  (3.204e-04 * sqft_living) + (1.164e-02 * yr_built) + 2.672e-04(sqft_living15) - 1.745e+01

Moving forward to predict the price our equation will be:
price = (6.200e+04 * grade) + (6.597e+01 * sqft_living) + (3.784e+01* sqft_living15) -2.076e+05

Plotting 3d model for multiple regression

```{r}

scatterplot3d(data2$price, data2$sqft_living, data2$sqft_living15, 
              color = data2$grade,
              type="p",)
```

















                                    
